---
title: Cherry Blossom Prediction 
subtitle: Exploratory Data Analysis (EDA)
author: Alex Yinan Guo
output:
  pdf_document:
    toc: true
    toc_depth: 4
---

```{r, include=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE) 
```

# Setup

Load packages and read in data sets.

```{r setup, results=FALSE}
library(GGally)
library(rjson)
library(tidyverse)
library(lubridate)
library(cowplot)
library(ggpubr)
library(knitr)
library(mice)
library(mctest)
library(lme4)
library(quantreg)
library(AER)
library(MASS)
library(dplyr)
library(reshape2)

df <- read.csv(file = '../data/processed/clean_data.csv')
df_non_impute <- read.csv(file = '../data/processed/processed_data.csv')
```

# Data Summaries

## Before Imputation

```{r}
summary(df_non_impute)
apply(df_non_impute, 2, function(x) length(unique(x)))
```

## Data After Imputation

```{r}
head(df)
tail(df)
summary(df)
apply(df, 2, function(x) length(unique(x)))
unique(df$country)
```



# Density Plots

## Overall Density Plots

```{r}
df %>%
  select_if(is.numeric) %>%
  dplyr::select(-year) %>%
  pivot_longer(everything()) %>%
  ggplot(aes(x=value)) + 
  geom_density(fill='grey') +
  facet_wrap(~name, scales='free') + 
  theme(strip.text = element_text(size=7),
        axis.text.x = element_text(size=5), 
        axis.text.y = element_text(size=5), 
        plot.title = element_text(hjust = 0.5)) +
  labs(x ="", 
       y = "Density",
       title = 'Overall Density Plots') 
```

## Density Plots of Japan

```{r}
df %>%
  filter(country=='Japan') %>%
  select_if(is.numeric) %>%
  dplyr::select(-year) %>%
  pivot_longer(everything()) %>%
  ggplot(aes(x=value)) + 
  geom_density(fill='grey') +
  facet_wrap(~name, scales='free') + 
  theme(strip.text = element_text(size=7),
        axis.text.x = element_text(size=5), 
        axis.text.y = element_text(size=5), 
        plot.title = element_text(hjust = 0.5)) +
  labs(x ="", 
       y = "Density",
       title = 'Density Plots of Japan') 
```

## Density Plots of South Korea

```{r}
df %>%
  filter(country=='South Korea') %>%
  select_if(is.numeric) %>%
  dplyr::select(-year) %>%
  pivot_longer(everything()) %>%
  ggplot(aes(x=value)) + 
  geom_density(fill='grey') +
  facet_wrap(~name, scales='free') + 
  theme(strip.text = element_text(size=7),
        axis.text.x = element_text(size=5), 
        axis.text.y = element_text(size=5), 
        plot.title = element_text(hjust = 0.5)) +
  labs(x ="", 
       y = "Density",
       title = 'Density Plots of South Korea') 
```

## Density Plots of Switzerland

```{r}
df %>%
  filter(country=='Switzerland') %>%
  select_if(is.numeric) %>%
  dplyr::select(-year) %>%
  pivot_longer(everything()) %>%
  ggplot(aes(x=value)) + 
  geom_density(fill='grey') +
  facet_wrap(~name, scales='free') + 
  theme(strip.text = element_text(size=7),
        axis.text.x = element_text(size=5), 
        axis.text.y = element_text(size=5), 
        plot.title = element_text(hjust = 0.5)) +
  labs(x ="", 
       y = "Density",
       title = 'Density Plots of Switzerland') 
```


## Density Plots of United States

```{r}
df %>%
  filter(country=='United States') %>%
  select_if(is.numeric) %>%
  dplyr::select(-year) %>%
  pivot_longer(everything()) %>%
  ggplot(aes(x=value)) + 
  geom_density(fill='grey') +
  facet_wrap(~name, scales='free') + 
  theme(strip.text = element_text(size=7),
        axis.text.x = element_text(size=5), 
        axis.text.y = element_text(size=5), 
        plot.title = element_text(hjust = 0.5)) +
  labs(x ="", 
       y = "Density",
       title = 'Density Plots of United States') 
```

# Year v.s. Numeric Variables 

## Year v.s. Numeric Variables of Japan

```{r}
par(mfrow=c(3, 4))

df %>%
  filter(country=='Japan') %>%
  select_if(is.numeric) %>%
  pivot_longer(-c(year, lat, long, alt)) %>%
  ggplot(aes(x=year, y=value)) + 
  geom_point(alpha = 0.05) +
  facet_wrap(~name, scales='free') +
  labs(x ="", 
       y = "",
       title="Year v.s. Numeric Variables of Japan") + 
  theme_light() +
  theme(strip.text = element_text(size=7),
        axis.text.x = element_text(size=7), 
        axis.text.y = element_text(size=7),
        plot.title = element_text(hjust = 0.5))
```

## Year v.s. Numeric Variables of South Korea

```{r}
par(mfrow=c(3, 4))

df %>%
  filter(country=='South Korea') %>%
  select_if(is.numeric) %>%
  pivot_longer(-c(year, lat, long, alt)) %>%
  ggplot(aes(x=year, y=value)) + 
  geom_point(alpha = 0.1) +
  facet_wrap(~name, scales='free') +
  labs(x ="", 
       y = "",
       title="Year v.s. Numeric Variables of South Korea") + 
  theme_light() +
  theme(strip.text = element_text(size=7),
        axis.text.x = element_text(size=7), 
        axis.text.y = element_text(size=7),
        plot.title = element_text(hjust = 0.5))
```

## Year v.s. Numeric Variables of Switzerland

```{r}
par(mfrow=c(3, 4))

df %>%
  filter(country=='Switzerland') %>%
  select_if(is.numeric) %>%
  pivot_longer(-c(year, lat, long, alt)) %>%
  ggplot(aes(x=year, y=value)) + 
  geom_point(alpha = 0.1) +
  facet_wrap(~name, scales='free') +
  labs(x ="", 
       y = "",
       title="Year v.s. Numeric Variables of Switzerland") + 
  theme_light() +
  theme(strip.text = element_text(size=7),
        axis.text.x = element_text(size=7), 
        axis.text.y = element_text(size=7),
        plot.title = element_text(hjust = 0.5))
```

## Year v.s. Numeric Variables of United States

```{r}
par(mfrow=c(3, 4))

df %>%
  filter(country=='United States') %>%
  select_if(is.numeric) %>%
  pivot_longer(-c(year, lat, long, alt)) %>%
  ggplot(aes(x=year, y=value)) + 
  geom_point(alpha = 0.1) +
  facet_wrap(~name, scales='free') +
  labs(x ="", 
       y = "",
       title="Year v.s. Numeric Variables of United States") + 
  theme_light() +
  theme(strip.text = element_text(size=7),
        axis.text.x = element_text(size=7), 
        axis.text.y = element_text(size=7),
        plot.title = element_text(hjust = 0.5))
```

# Visualization of Data Locations

```{r locations, warning=FALSE, fig.align = "center"}
world <- map_data("world")
all_plot <- ggplot() +
  geom_map(
    data = world, map = world,
    aes(long, lat, map_id = region, fill = region),
    color = "gray", fill = "white") +
  geom_point(
    data = df,
    aes(long, lat, color=country), 
    alpha=0.1, size=1) + 
  guides(colour = guide_legend(override.aes = list(alpha=0.9,size=4))) + 
  labs(title='Data Locations') + 
  theme(plot.title = element_text(hjust = 0.5))

jp <- map_data('world', 'Japan')
jp_plot <- ggplot() +
  geom_map(
    data = jp, map = jp,
    aes(long, lat, map_id = region),
    color = "white", fill = "white"
  ) +
  geom_point(
    data = df %>% filter(country=='Japan'),
    aes(long, lat), alpha=0.01) + 
  ggtitle("Japan") + 
  theme(plot.title = element_text(hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5))

swiss <- map_data('world', 'Switzerland')
swiss_plot <- ggplot() +
  geom_map(
    data = swiss, map = swiss,
    aes(long, lat, map_id = region),
    color = "gray", fill = "white"
  ) +
  geom_point(
    data = df %>% filter(country=='Switzerland'),
    aes(long, lat), alpha=0.01) + 
  ggtitle("Switzerland") + 
  theme(plot.title = element_text(hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5))

sk <- map_data('world', 'South Korea')
sk_plot <- ggplot() +
  geom_map(
    data = sk, map = sk,
    aes(long, lat, map_id = region),
    color = "white", fill = "white"
  ) +
  geom_point(
    data = df %>% filter(country=='South Korea'),
    aes(long, lat), alpha=0.01) + 
  ggtitle("South Korea") + 
  theme(plot.title = element_text(hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5))

usa <- map_data("state")
usa_plot <- ggplot() +
  geom_map(
    data = usa, map = usa,
    aes(long, lat, map_id = region),
    fill = "white", size = 0.1
  ) +
  geom_point(
    data = df %>% filter(country=='United States'),
    aes(long, lat), alpha=0.1) + 
  ggtitle("U.S.") + 
  theme(plot.title = element_text(hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5))

individual_plots <- ggarrange(jp_plot, swiss_plot, sk_plot, usa_plot, 
          ncol = 4, nrow = 1)

ggarrange(all_plot, individual_plots, 
          ncol = 1, nrow = 2, heights = c(2.5,1.5))
```

# Basic Linear Models

```{r}
df <- read.csv("../data/processed/clean_data.csv")
summary(df)

df <- df %>% na.omit
train <- df[sort(sample(nrow(df), nrow(df)*.7)),]
test <- df[-sort(sample(nrow(df), nrow(df)*.7)),]
```

## Fit Linear Regression

```{r}
lm <- lm(bloom_doy~.-country-city-year, data = train) 
summary(lm)

plot(predict(lm, test), 
     test$bloom_doy, 
     xlab = "Predicted Values", 
     ylab = "Observed Values", 
     main = 'Predicted Values v.s. Observed Values of Linear Regression')
abline(a=0, b=1, lib = 1, col = "red", lwd = 2)

plot(lm, which = c(1,2,3,4,5)) 
```

## Confidence Intervals

The confidence interval reflects the uncertainty around the mean predictions. To display the 95% confidence intervals around the mean the predictions, specify the option interval = "confidence".

For example, the 95% confidence interval associated with the first observation in the test data set is (43.02558, 47.14276). This means that, according to our model, a tree with same parameters with the first observation has, on average, a bloom_doy between 43 and 47.

```{r}
lm_confidence_intervals <- predict(lm, newdata = test, interval = "confidence")
head(lm_confidence_intervals)
```

## Prediction Intervals

The prediction interval gives uncertainty around a single value. In the same way, as the confidence intervals, the prediction intervals can be computed as follow.

The 95% prediction intervals associated with the first observation in the test dataset is (27.33479, 62.83354). This means that, according to our model, 95% of trees with same parameters have bloom_doy between 27 and 63.

```{r}
lm_prediction_intervals <- predict(lm, newdata = test, interval = "prediction")
head(lm_prediction_intervals)
```

## Check multicollinearity 

First we check the correlation matrix of independent variables. We found that there are serveral variables that have high correlation with each other. 

```{r}
X <- df %>% dplyr::select(-c(country, city, year, bloom_doy))
cormat <- round(cor(X),2)

reorder_cormat <- function(cormat){
  # Use correlation between variables as distance
  dd <- as.dist((1-cormat)/2)
  hc <- hclust(dd)
  cormat <-cormat[hc$order, hc$order]
}

# Get lower triangle of the correlation matrix
get_lower_tri<-function(cormat){
    cormat[upper.tri(cormat)] <- NA
    return(cormat)
}

# Get upper triangle of the correlation matrix
get_upper_tri <- function(cormat){
    cormat[lower.tri(cormat)]<- NA
    return(cormat)
}

# Reorder the correlation matrix
cormat <- reorder_cormat(cormat)
upper_tri <- get_upper_tri(cormat)
# Melt the correlation matrix
melted_cormat <- melt(upper_tri, na.rm = TRUE)
# Create a ggheatmap
ggheatmap <- ggplot(melted_cormat, aes(Var2, Var1, fill = value))+
 geom_tile(color = "white")+
 scale_fill_gradient2(low = "blue", high = "red", mid = "white", 
   midpoint = 0, limit = c(-1,1), space = "Lab", 
    name="Pearson\nCorrelation") +
  theme_minimal()+ # minimal theme
 theme(axis.text.x = element_text(angle = 45, vjust = 1, 
    size = 12, hjust = 1))+
 coord_fixed() + 
geom_text(aes(Var2, Var1, label = value), color = "black", size = 2) +
theme(
  axis.title.x = element_blank(),
  axis.title.y = element_blank(),
  panel.grid.major = element_blank(),
  panel.border = element_blank(),
  panel.background = element_blank(),
  axis.ticks = element_blank(),
  legend.justification = c(1, 0),
  legend.position = c(0.6, 0.7),
  legend.direction = "horizontal")+
  guides(fill = guide_colorbar(barwidth = 7, barheight = 1,
                title.position = "top", title.hjust = 0.5))
# Print the heatmap
print(ggheatmap)
```


The ‘mctest’ package in R provides the Farrar-Glauber test and other relevant tests for multicollinearity. Here we check the overall and individual diagnostic checking for multicollinearity respectively. The calculated value of the Chi-square test statistic is found to be 49102.7, implying the presence of multicollinearity in the model. 

Next, we check Farrar–Glauber test (F-test) for the location of the multicollinearity. The VIF, TOL and Wi columns provide the diagnostic output for variance inflation factor, tolerance and Farrar-Glauber F-test respectively. Variables such as `tmin`, `tmin_winter`, and `tmax_winter` show high correlation with other features.

```{r}
options(scipen=999)
omcdiag(lm)
imcdiag(lm)
```


## Mix Effects Model

```{r}
suppressWarnings(suppressMessages(
  print(full_mixed_model <- lmer(bloom_doy ~ tmin_winter+
                                   prcp +
                                   agdd_winter +
                                   prcp_winter +
                                   co2_percapita + 
                                   (tmin_winter +
                                      prcp +
                                      agdd_winter +
                                      prcp_winter +
                                      co2_percapita | city), 
                                 data = train
                                 ))))
summary(full_mixed_model)
head(coef(full_mixed_model)$city)
```

```{r}
rq <- rq(bloom_doy~tmin_winter+prcp+agdd_winter+prcp_winter+co2_percapita, 
         data = train, 
         tau=c(0.25, 0.5, 0.75))
summary(rq)
plot(rq)

head(round(predict(rq, newdata = test), 0))
```

