---
title: "Cherry Blossom Prediction"
author: "Amelia Tang, Alex Yinan Guo, Nick Lisheng Mao"
date: '2022/02/23 (updated: `r Sys.Date()`)'
output:
  html_document:
    toc: yes
  github_document:
    toc: yes
always_allow_html: yes
bibliography: cherry_blossom_reference.bib
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(knitr)
library(kableExtra)
library(tidyverse)
library(scales)
```

```{r read_test_score, include=FALSE}
#test_score <- read_csv("xxx") 
``` 

## Summary

TBU (Amelia)

## Introduction

Amelia

## Methods
### Data Collection and Processing 

Data Collection (Nick)
Processing (Nick)
Imputation (Amelia)

### Exploratory Data Analysis

TBU - (Alex) 
For the weather data - treat as time series 
For boy - regression 

```{r target, fig.align = 'center', echo=FALSE, fig.cap="Figure 1. Target Distribution", out.width = '40%'}
#knitr::include_graphics("xxx")
```

### Model Selection
- Linear regression (Alex)
- Tree based (Alex)
- distance-based (Amelia)
- Unsupervised Kmeans + SVR (Amelia)
- Hyper parameter tuning (Alex)

### Forecasting 
Weather data (nick)

## Results & Discussion
results (Amelia)
discussion (limitations - more data, try different algorithms etc.)

# References
TBU 