---
title: Cherry Blossom Prediction 
subtitle: Exploratory Data Analysis (EDA)
author: Alex Yinan Guo
output:
  pdf_document:
    toc: true
    toc_depth: 4
---

```{r, include=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE) 
```

# Setup

Load packages and read in data sets.

```{r setup, results=FALSE}
library(GGally)
library(rjson)
library(tidyverse)
library(lubridate)
library(cowplot)
library(ggpubr)

washingtondc <- read.csv(file = '../data/raw/washingtondc.csv')
south_korea <- read.csv(file = '../data/raw/south_korea.csv')
kyoto <- read.csv(file = '../data/raw/kyoto.csv')
japan <- read.csv(file = '../data/raw/japan.csv')
liestal <- read.csv(file = '../data/raw/liestal.csv')
meteoswiss <- read.csv(file = '../data/raw/meteoswiss.csv')
USA_status_intensity <- 
  read.csv(file = '../data/raw/USA-NPN_status_intensity_observations_data.csv') %>%
  filter(!State %in% c('Okinawa', 'Liaoning Sheng', 'Hyogo', 'Chongqing Shi')) %>%
  mutate(Observation_Date = mdy(Observation_Date),
         Year = format(Observation_Date, format = "%Y"),
         Date = format(Observation_Date, format = "%m-%d"))
USA_individual_phenometrics <- 
  read.csv(file = '../data/raw/USA-NPN_individual_phenometrics_data.csv') %>%
  filter(!State %in% c('Okinawa', 'Liaoning Sheng', 'Hyogo', 'Chongqing Shi'))
```

# Summaries of Data Sets

```{r summary}
summary(washingtondc)
summary(liestal)
summary(kyoto)

summary(south_korea)
unique(south_korea$location)
summary(japan)
unique(japan$location)
summary(meteoswiss)
unique(meteoswiss$location)

summary(USA_status_intensity)

summary(USA_individual_phenometrics)
# Species in the U.S.
unique(USA_individual_phenometrics%>%select(Common_Name))
# Species in Washington, D.C.
unique(USA_individual_phenometrics%>%filter(State=='DC')%>%select(Common_Name))
```


# Exploration of City Data Sets

For this competition, we are going to forecast the cherry blossom peak of the following four cities:

- Washington, D.C. (USA)
- Kyoto (Japan)
- Liestal-Weideli (Switzerland) 
- Vancouver, BC (Canada) 

## Locations of Cities

```{r cities, warning=FALSE}
world <- map_data("world")
cities <- rbind(kyoto[1,],washingtondc[1,],liestal[1,]) %>% 
  select(lat, long) %>% 
  add_row(lat=49.246292, long=-123.116226) %>%
  add_column(location=c('Kyoto','Washington DC', 'Liestal', 'Vancouver'))

ggplot() +
  geom_map(
    data = world, 
    map = world,
    aes(long, lat, map_id = region),
    color = "gray", fill = "white") +
  geom_point(
    data = cities,
    aes(long, lat, color=location), size=3) 
```

## Time Series Visualization of Peak Bloom Dates

We use the provided historical bloom data of three cities where the cherry blossom peak bloom dates are going to be predicted to plot time series.

```{r Year v.s. bloom day plots, warning=FALSE, fig.height = 2, fig.width = 8, fig.align = "center"}
ggplot() + 
  geom_line(data = washingtondc, 
            aes(year, bloom_doy, color='Washington DC, US'), 
            alpha=0.7, 
            size=0.6) + 
  geom_line(data = kyoto %>% filter(year>1900), 
            aes(year, bloom_doy, color='Kyoto, Japan'), 
            alpha=0.7, 
            size=0.6) +
  geom_line(data = meteoswiss %>% 
              filter(location=='Switzerland/Liestal', year>1980), 
            aes(year, bloom_doy, color='Liestal, Switzerland'), 
            alpha=0.7, 
            size=0.6) + 
  labs(title="Peak Bloom Dates 1900 - 2021",
       subtitle = "Washington DC, Kyoto, Liestal",
       x ="Year", 
       y = "Day of the peak bloom",
       color = "City") +
  scale_y_discrete(limit = seq(85, 125, 10),
                   labels = c('March 26', 
                              'April 5', 
                              'April 15', 
                              'April 25', 
                              'May 5')) + 
  theme(plot.title = element_text(hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5))
```


# Exploration of Country Data Sets

Here we explore more about the following data sets containing information regarding cherry blossom of different locations in four countries:

- south_korea.csv
- japan.csv
- meteoswiss.csv
- USA-NPN_status_intensity_observations_data.csv

## Visualization of Data Locations

```{r locations, warning=FALSE}
ggplot() +
  geom_map(
    data = world, map = world,
    aes(long, lat, map_id = region, fill = region),
    color = "gray", fill = "white") +
  geom_point(
    data = rbind(japan %>% mutate(location=sub("/.*", "",location)) %>% 
                   select(location,long,lat), 
                 meteoswiss %>% mutate(location=sub("/.*", "",location)) %>% 
                   select(location,long,lat),
                 south_korea %>% mutate(location=sub("/.*", "",location)) %>% 
                   select(location,long,lat),
                 USA_status_intensity %>%  
                   select(Longitude, Latitude) %>%
                   mutate(location='the U.S.') %>%
                   rename(long=Longitude, lat=Latitude)),
    aes(long, lat, color=location), 
    alpha=0.1, size=0.1) + 
  guides(colour = guide_legend(override.aes = list(alpha=0.9,size=4)))
```

```{r jp swiss sk usa locations, warning=FALSE, fig.height = 2, fig.width = 10, fig.align = "center"}
jp <- map_data('world', 'Japan')
jp_plot <- ggplot() +
  geom_map(
    data = jp, map = jp,
    aes(long, lat, map_id = region),
    color = "white", fill = "white"
  ) +
  geom_point(
    data = japan,
    aes(long, lat), alpha=0.01) + 
  ggtitle("Sites in Japan") + 
  theme(plot.title = element_text(hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5))

swiss <- map_data('world', 'Switzerland')
swiss_plot <- ggplot() +
  geom_map(
    data = swiss, map = swiss,
    aes(long, lat, map_id = region),
    color = "gray", fill = "white"
  ) +
  geom_point(
    data = meteoswiss,
    aes(long, lat), alpha=0.01) + 
  ggtitle("Sites in Switzerland") + 
  theme(plot.title = element_text(hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5))

sk <- map_data('world', 'South Korea')
sk_plot <- ggplot() +
  geom_map(
    data = sk, map = sk,
    aes(long, lat, map_id = region),
    color = "white", fill = "white"
  ) +
  geom_point(
    data = south_korea,
    aes(long, lat), alpha=0.01) + 
  ggtitle("Sites in South Korea") + 
  theme(plot.title = element_text(hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5))

usa <- map_data("state")
usa_plot <- ggplot() +
  geom_map(
    data = usa, map = usa,
    aes(long, lat, map_id = region),
    fill = "white", size = 0.1
  ) +
  geom_point(
    data = USA_status_intensity,
    aes(Longitude, Latitude), alpha=0.1) + 
  ggtitle("Sites in the U.S.") + 
  theme(plot.title = element_text(hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5))

ggarrange(jp_plot, swiss_plot, sk_plot, usa_plot, 
          ncol = 4, nrow = 1)
```
## Density Plots

```{r}
japan %>%
  select_if(is.numeric) %>%
  select(-year) %>%
  pivot_longer(everything()) %>%
  ggplot(aes(x=value)) + 
  geom_density(fill='grey') +
  facet_wrap(~name, scales='free') + 
  theme(strip.text = element_text(size=7),
        axis.text.x = element_text(size=5), 
        axis.text.y = element_text(size=5), 
        plot.title = element_text(hjust = 0.5)) +
  labs(x ="", 
       y = "Density",
       title = 'Density Plots of Japan') 
```

```{r}
south_korea %>%
  select_if(is.numeric) %>%
  select(-year) %>%
  pivot_longer(everything()) %>%
  ggplot(aes(x=value)) + 
  geom_density(fill='grey') +
  facet_wrap(~name, scales='free') + 
  theme(strip.text = element_text(size=7),
        axis.text.x = element_text(size=5), 
        axis.text.y = element_text(size=5), 
        plot.title = element_text(hjust = 0.5)) +
  labs(x ="", 
       y = "Density",
       title = 'Density Plots of South Korea') 
```

```{r}
meteoswiss %>%
  select_if(is.numeric) %>%
  select(-year) %>%
  pivot_longer(everything()) %>%
  ggplot(aes(x=value)) + 
  geom_density(fill='grey') +
  facet_wrap(~name, scales='free') + 
  theme(strip.text = element_text(size=7),
        axis.text.x = element_text(size=5), 
        axis.text.y = element_text(size=5), 
        plot.title = element_text(hjust = 0.5)) +
  labs(x ="", 
       y = "Density",
       title = 'Density Plots of Switzerland') 
```

```{r}
USA_status_intensity %>%
  select_if(is.numeric) %>%
  select(-c(Abundance_Value, 
            Individual_ID, 
            Intensity_Category_ID, 
            Observation_ID, 
            Phenophase_ID, 
            Phenophase_Status, 
            Site_ID, 
            Species_ID)) %>%
  filter(Prcp_Spring>0, 
         Prcp_Winter>0,
         Tmax_Spring != '-9999',
         Tmax_Winter != '-9999',) %>%
  pivot_longer(everything()) %>%
  ggplot(aes(x=value)) + 
  geom_density(fill='grey') +
  facet_wrap(~name, scales='free') + 
  theme(strip.text = element_text(size=7),
        axis.text.x = element_text(size=5), 
        axis.text.y = element_text(size=5), 
        plot.title = element_text(hjust = 0.5)) +
  labs(x ="", 
       y = "Density",
       title = 'Density Plots of United States') 
```


## Time Series Visualization of Cities

Here we randomly selected some cities in each data set to plot the peak bloom date against year. We only create line charts for cities in Japan, Switzerland, and South Korea. The number of usable observations are too low in the U.S. data set, so we will not visualize it.

```{r jp sk cities plot, warning=FALSE, fig.align = "center"}
japan_plot <- japan %>% mutate(location=gsub("Japan/","",location)) %>% 
  filter(location==c('Nagoya', 'Tokyo', 'Kyoto', 'Osaka'))%>%
  ggplot(aes(year, bloom_doy)) +
  aes(x = year, color=location) +
  geom_line() +
  labs(title="Japan Bloom Day Time Series",
       x =" ", y = " ") +
  scale_y_discrete(limit = seq(85, 105, 10),
                   labels = c('March 26', 'April 5', 'April 15')) + 
  theme(plot.title = element_text(hjust = 0.5))

meteoswiss_plot <- meteoswiss %>% 
  mutate(location=gsub("Switzerland/","",location)) %>% 
  arrange(location) %>% 
  filter(location==c('Changins', 'Basel-Binningen', 'Cartigny', 'Versoix')) %>%
  ggplot(aes(year, bloom_doy)) +
  aes(x = year, color=location) +
  geom_line() +
  labs(title="Switzerland Bloom Day Time Series",
       x =" ", y = "Day of the bloom") +
  scale_y_discrete(limit = seq(85, 105, 10),
                   labels = c('March 26', 'April 5', 'April 15')) + 
  theme(plot.title = element_text(hjust = 0.5))

south_korea_plot <- south_korea %>% 
  mutate(location=gsub("South Korea/","",location)) %>% 
  arrange(location) %>% 
  filter(location==c('Seoul', 'Incheon', 'Busan', 'Jeju')) %>%
ggplot(aes(year, bloom_doy)) +
    aes(x = year, color=location) +
    geom_line(na.rm = FALSE) +
  labs(title="South Korea Bloom Day Time Series",
        x ="Year", y = " ") +
  scale_y_discrete(limit = seq(85, 105, 10),
                   labels = c('March 26', 'April 5', 'April 15')) + 
  theme(plot.title = element_text(hjust = 0.5))

ggarrange(japan_plot, meteoswiss_plot,south_korea_plot,
          ncol = 1, nrow = 3)
```

## Wxploration of USA NPN Data Sets

### Relationships Between AGDD and Peak Bloom Dates

Before visualization, we filter the data by intensity value over 75% because we want to explore more about relationships between peak bloom date and other variables. 

```{r}
USA_status_intensity <- USA_status_intensity %>%
  filter(Intensity_Value=='75-94%') %>%
  group_by(Year, Site_ID) %>% 
  slice(which.min(Day_of_Year))

USA_status_intensity %>%
filter(AGDD > 0) %>%
  ggplot() + 
  geom_point(aes(x = Observation_Date, y = AGDD)) +
  labs(title="US AGDD v.s. Peak Bloom Date",
        x ="Date", y = "AGDD (celcius)") + 
  theme(plot.title = element_text(hjust = 0.5), 
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
```



```{r}
USA_status_intensity %>%
  filter(AGDD > 0) %>%
  ggplot() + 
  geom_point(aes(x = Date, y = AGDD)) +
  facet_wrap(~Year) +
  labs(title="US AGDD v.s. Peak Bloom Dates by Year",
        x ="Date", y = "AGDD (celcius)") + 
  scale_x_discrete(breaks = c('04-10','04-20','05-20')) +
  theme(plot.title = element_text(hjust = 0.5), 
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
```

### Relationships Between Temperature and Peak Bloom Dates

```{r}
USA_status_intensity %>%
  filter(Tmin_Winter !=-9999) %>%
  ggplot() + 
  geom_point(aes(x = Date, y = Tmax_Spring, color='Tmax_Spring')) +
  geom_point(aes(x = Date, y = Tmin_Spring, color='Tmin_Spring')) +
#  geom_point(aes(x = Date, y = Tmax_Winter, color='Tmax_Winter')) +
#  geom_point(aes(x = Date, y = Tmin_Winter, color='Tmin_Winter')) +
  facet_wrap(~Year) +
  labs(title="US Spring Temperature v.s. Peak Bloom Dates",
        x ="Date", y = "Temperature", color='Year') + 
  scale_x_discrete(breaks = c('04-10','05-01','05-20')) +
  theme(plot.title = element_text(hjust = 0.5), 
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
```

```{r}
USA_status_intensity %>%
  filter(Tmin_Winter !=-9999) %>%
  ggplot() + 
  geom_point(aes(x = Date, y = Tmax_Spring, color='Tmax_Spring'), alpha=0.5) +
  geom_point(aes(x = Date, y = Tmin_Spring, color='Tmin_Spring'), alpha=0.5) +
  geom_point(aes(x = Date, y = Tmax_Winter, color='Tmax_Winter'), alpha=0.5) +
  geom_point(aes(x = Date, y = Tmin_Winter, color='Tmin_Winter'), alpha=0.5) +
  facet_wrap(~Year) +
  labs(title="US Temperature v.s. Peak Bloom Dates by Year",
        x ="Date", y = "Temperature", color='Year') + 
  scale_x_discrete(breaks = c('04-10','05-01','05-20')) +
  theme(plot.title = element_text(hjust = 0.5), 
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
```
### Relationships Between Pricipitation and Peak Bloom Dates

```{r}
USA_status_intensity %>%
  filter(Prcp_Spring > 0, Prcp_Winter > 0) %>%
  ggplot() + 
  geom_point(aes(x = Observation_Date, y = Prcp_Winter, color='Winter')) +
  geom_point(aes(x = Observation_Date, y = Prcp_Spring, color='Spring')) +
  labs(title="US Pricipitation v.s. Peak Bloom Date ",
        x ="Date", y = "Pricipitation") + 
  theme(plot.title = element_text(hjust = 0.5), 
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
```

```{r}
USA_status_intensity %>%
  filter(Prcp_Spring > 0, Prcp_Winter > 0) %>%
  ggplot() + 
  geom_point(aes(x = Date, y = Prcp_Spring, color = 'Spring')) +
  geom_point(aes(x = Date, y = Prcp_Winter, color = 'Winter')) +
  facet_wrap(~Year) +
  labs(title="US Precipitation v.s. Peak Bloom Date by Year",
        x ="Date", y = "Precipitation", color='Season') + 
  scale_x_discrete(breaks = c('04-10','05-01','05-20')) +
  theme(plot.title = element_text(hjust = 0.5), 
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
```




